---
title: |
    Suggested TMT processing pipeline for PNNL DMS phospho data.

vignette: >
  %\VignetteIndexEntry{tmt_pipeline_v1_PNNL_DMS_phospho}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

output:
  BiocStyle::html_document:
     toc: true
---


```{r setup, echo=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
# knitr::opts_chunk$set(echo=T, message=F, warning=F, fig.align='center', out.width='10cm')
```


`PlexedPiper` is the main library for the pipeline. `MSnID` package is a 
flexible way of handling MS/MS identifications. It handles MS/MS data
filtering part.

```{r libraries, eval=FALSE}
library(PlexedPiper)
library(MSnID)
library(tidyverse)
```


# Aspects of PNNL DMS

PNNL DMS is based on MS SQL Server. Typically data is organized into data packages according to experiments. Data package that contains data mirroring `PlexedPiperTestData` is number `3442`.

# Prepare MS/MS IDs

## Read the MS-GF+ output

First step is to determine MS-GF+ jobs using the data package number.


This simply reads parsed to text output of MS-GF+ search engine. The text files
are collated together and the resulting `data.frame` used to create MSnID object.
```{r msms, eval=FALSE}
# phospho
msnid <- read_msms_data_from_DMS(3432)
ascore <- get_AScore_results(3432)
msnid <- best_PTM_location_by_ascore(msnid, ascore)
# at some point I need to filter non-phospho away
msnid <- filter_msgf_data_peptide_level(msnid, 0.01)
msnid <- remap_accessions_uniprot_to_gene(msnid, organism_name="Homo sapiens")

path_to_FASTA <- path_to_FASTA_used_by_DMS(3432)
path_to_FASTA_gene <- remap_accessions_uniprot_to_gene_fasta(path_to_FASTA)

# todo. or done? seems to be working
msnid <- compute_num_peptides_per_1000aa(msnid, path_to_FASTA_gene)



show(msnid)
msnid <- filter_msgf_data_protein_level(msnid, 0.01)
show(msnid)

# no need for parsimonius inference with phospho
# msnid <- infer_parsimonious_accessions(msnid, unique_only=TRUE)
# show(msnid)

msnid <- apply_filter(msnid, "!isDecoy")



# remap to site notation
# needed: 1) accession, 
#         2) fasta with accession, 
#         3) peptide sequence with asterisc location
# Let's borrow vp.misc remapping. !!! Ultimately it should be moved to MSnID
fst <- readAAStringSet(path_to_FASTA_gene, format="fasta", 
                       nrec=-1L, skip=0L, use.names=TRUE)
ids <- psms(msnid) %>%
   distinct(accession, peptide)
ids_with_sites <- map_PTM_sites(ids, fst, "accession", "peptide", "*")

ids_with_sites <- ids_with_sites %>%
   mutate(idx = map(PepLoc, length) %>% unlist) %>%
   filter(idx != 0) %>%
   dplyr::select(-idx)

ids_with_sites_simplified <- ids_with_sites %>%
   dplyr::select(accession, peptide, SiteCollapsedFirst) %>%
   mutate(site = unlist(SiteCollapsedFirst) %>%
             gsub(",","_",.) %>%
             paste(accession, ., sep="-")) %>%
   dplyr::select(-SiteCollapsedFirst)

psms(msnid) <- inner_join(psms(msnid), ids_with_sites_simplified)

psms(msnid) <- psms(msnid) %>%
   inner_join(ids_with_sites_simplified)






masic_data <- read_masic_data_from_DMS(3432, interference_score = T)
masic_data <- filter_masic_data(masic_data, 0.5, 0)

library(dplyr)
fractions <- masic_data %>%
   distinct(Dataset) %>%
   mutate(PlexID = sub("SHSY5Y_IIS_3x3_P_(B\\d).*","\\1",Dataset))
head(fractions)

library(readr)
samples <- read_tsv("../../3432/samples.txt")
head(samples,10)

ref <- "(`10min_control1`*`10min_control2`*`10min_Insulin`*`10min_IGF1`*`10min_IGF2`*`60min_control1`*`60min_control2`*`60min_Insulin`*`60min_IGF1`*`60min_IGF2`)^(1/10)"
references <- samples %>%
   distinct(PlexID, QuantBlock) %>%
   mutate(Reference = ref)


aggregation_level <- c("site")
quant_cross_tab <- create_crosstab(msnid, 
                                   masic_data, 
                                   aggregation_level, 
                                   fractions, samples, references)
dim(quant_cross_tab)
head(quant_cross_tab)



```





