---
title: |
    Suggested TMT processing pipeline for PNNL DMS phospho data.

vignette: >
  %\VignetteIndexEntry{tmt_pipeline_v1_PNNL_DMS_phospho}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

output:
  BiocStyle::html_document:
     toc: true
---


```{r setup, echo=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
# knitr::opts_chunk$set(echo=T, message=F, warning=F, fig.align='center', out.width='10cm')
```


`PlexedPiper` is the main library for the pipeline. `MSnID` package is a 
flexible way of handling MS/MS identifications. It handles MS/MS data
filtering part.

```{r libraries, eval=FALSE}
library(PlexedPiper)
library(MSnID)
library(tidyverse)
```


# Aspects of PNNL DMS

PNNL DMS is based on MS SQL Server. Typically data is organized into data packages according to experiments. Data package that contains data mirroring `PlexedPiperTestData` is number `3442`.

# Prepare MS/MS IDs

## Read the MS-GF+ output

First step is to determine MS-GF+ jobs using the data package number.


This simply reads parsed to text output of MS-GF+ search engine. The text files
are collated together and the resulting `data.frame` used to create MSnID object.
```{r msms, eval=FALSE}
# phospho
msnid <- read_msms_data_from_DMS(3432)
ascore <- get_AScore_results(3432)
msnid <- best_PTM_location_by_ascore(msnid, ascore)
msnid <- filter_msgf_data_peptide_level(msnid, 0.01)
msnid <- remap_accessions_uniprot_to_gene(msnid, organism_name="Homo sapiens")

path_to_FASTA <- path_to_FASTA_used_by_DMS(3432)
path_to_FASTA_gene <- remap_accessions_uniprot_to_gene_fasta(path_to_FASTA)


# todo
msnid <- compute_num_peptides_per_1000aa(msnid, path_to_FASTA_gene)



show(msnid)
msnid <- filter_msgf_data_protein_level(msnid, 0.01)
show(msnid)

# msnid <- infer_parsimonious_accessions(msnid, unique_only=TRUE)
show(msnid)

msnid <- apply_filter(msnid, "!isDecoy")

masic_data <- read_masic_data_from_DMS(3432, interference_score = T)




```





