---
title: |
    Suggested TMT processing pipeline. Variant 1.

vignette: >
  %\VignetteIndexEntry{tmt_pipeline_v1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

output:
  BiocStyle::html_document:
     toc: true
---


```{r setup, echo=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
# knitr::opts_chunk$set(echo=T, message=F, warning=F, fig.align='center', out.width='10cm')
```


`PlexedPiper` is the main library for the pipeline. `MSnID` package is a 
flexible way of handling MS/MS identifications. It handles MS/MS data
filtering part.

```{r libraries}
library(PlexedPiper)
library(MSnID)
```


# Data source

This pipeline shows how to process TMT data that is processed outside of PNNL DMS.
This implies that raw files processed by MSGF+ and MASIC tools and the
processing results are in corresponding directories. For convienience the results
of MSGF+ and MASIC processing are provided in a companion `PlexedPiperTestData`
package.


# Prepare MS/MS IDs

## Read the MS-GF+ output

This simply reads parsed to text output of MS-GF+ search engine. The text files
are collated together and the resulting `data.frame` used to create MSnID object.
```{r msms}
path_to_MSGF_results <- system.file(
   "extdata/global/msgf_output",
   package = "PlexedPiperTestData")
msnid <- read_msgf_data(path_to_MSGF_results)
show(msnid)
```

This `MSnID` methods corrects for isotope selection error. Occasionally,
instrument selects not the lightest or monoisotopic peak, but a peak
with +1 or more C13. While MS-GF+ is still capable of correctly identifying 
those, the downstream calculations of mass measurement error need to be fixed.
This method corrects mass measurement errors accounting for selection of
other than monoisotopic peak.
```{r correct_isotopes}
msnid <- correct_peak_selection(msnid)
```

## MS/MS ID filter and peptide level

Main MS/MS filtering step. The `filter_msgf_data_peptide_level` function
wraps-up `MSnID::optimize_filter` method and uses `PepQValue` from MS-GF+ output
and absolute deviation of mass measurement error of parent ions (in ppm).
```{r peptide_level_filter}
show(msnid)
msnid <- filter_msgf_data_peptide_level(msnid, 0.01)
show(msnid)
```

## Switching annotation from RefSeq to gene symbols

This RefSeq to gene symbols conversion step can be skipped if the downstream
analysis relies on RefSeq IDs.

RefSeq is a very redundant database and contains lots of tentative splice 
isoforms and unconfirmed ORFs. As a result, a lot of peptides (~60%) are shared
across multiple RefSeq accessions. This adds to uncertainity which exactly 
protein we are quantifying. One solution is to switch to less redundant 
annotations like gene symbol. Moreover gene symbols are more human-interpretable
and quite commonly used in downstream analysis.
```{r refseq_2_gene}
msnid <- remap_accessions_refseq_to_gene(msnid, 
                                         organism_name="Rattus norvegicus")
```

This step converts sequences names in the fasta file used for MS/MS searches 
from RefSeq to gene symbols. This is necessary if we are going to count the
number of peptide identifications per protein or protein sequence coverage with
accessions switched from RefSeq to gene symbols.
```{r fasta_refseq_2_gene}
path_to_FASTA <- system.file(
   "extdata/Rattus_norvegicus_NCBI_RefSeq_2018-04-10.fasta.gz",
   package = "PlexedPiperTestData")
temp_dir <- tempdir()
file.copy(path_to_FASTA, temp_dir)
path_to_FASTA <- file.path(temp_dir, basename(path_to_FASTA))
path_to_FASTA_gene <- remap_accessions_refseq_to_gene_fasta(
   path_to_FASTA, organism_name="Rattus norvegicus")
```


## MS/MS ID filter at protein level

A while ago proteomics field established hard-and-fast two peptide per protein
rule. That is if a protein identified with one peptide it is not considered a
confident identification. It works OK, but there is a room for improvement. 
For example this rule really penalizes short protens. Then there are some very
long proteins (e.g. Titin 3.8 MDa) that easily have more then two matching 
peptides even in reversed sequence. Thus, here we propose to normalize number
of peptides per protein length and use that as a filtering criterion.
```{r protein_level_filter}
msnid <- compute_num_peptides_per_1000aa(msnid, path_to_FASTA_gene)
show(msnid)
msnid <- filter_msgf_data_protein_level(msnid, 0.01)
show(msnid)
```


## Inference of parsimonious protein set

The situation when sequence of a certain peptide matches multiple proteins adds
complication to the downstream quantitative analysis as it is not clear which 
protein product this peptide is originating from. There are common ways for
selecting a protein set. One is simply retain only uniquely matching peptides
(`unique_only=TRUE`). Shared peptides are simply discarded in this case.
Alternatively (in case of `unique_only=FALSE`) the shared peptides assigned to
the proteins with larger number of uniquely mapping peptides. If there is a 
choice between multiple proteins with equal number of uniquely mapping peptides,
the shared peptides is assigned to the first protein according to alphanumeric
order.

```{r inference}
msnid <- infer_parsimonious_accessions(msnid, unique_only=TRUE)
show(msnid)
```


## Remove decoy accessions

```{r remove_decoy}
msnid <- apply_filter(msnid, "!isDecoy")
```


# Prepare reporter ion intensities

## Read MASIC ouput

MASIC is a tool for extracting ion intensities. With proper parameter settings
it can be used for extracting TMT (or iTRAQ) reporter ion intensities. 
In addition it reporter a number of other helpful metrics. Of the most interest
are interference score of the parent ion level and S/N at reporter ion level.
Interference score reflects what proportion of ion population isolated for 
fragmentation is due to the targeted ion. The other words 1 - InterferenceScore
is due to co-isolated species that have similar elution time and parent ion m/z.
S/N at reporter level is the metric that computed by Thermo software. 

```{r read_masic}
path_to_MASIC_results <- system.file("extdata/global/masic_output", package = "PlexedPiperTestData")
masic_data <- read_masic_data(path_to_MASIC_results, interference_score=TRUE)
head(masic_data)
```


## Filtering MASIC data

In this case (currently recommended filters) we require that at least 50% of
ion population is due to the targeted ion and no filter at S/N level.

```{r filter_masic}
nrow(masic_data)
masic_data <- filter_masic_data(masic_data, 0.5, 0)
nrow(masic_data)
```



# Creating cross-tab

## Laying out study desing
To convert from PSMs and reporter ion intensities to meaningful quantitative
data it is necessary to know what are the samples in the reporter channels and
what is the intended reference channel (or combination of channels). The entire
study desing captured by three tables - fractions, samples, references.

## Fractions
This links dataset to original plex.
```{r read_fractions}
library(readr)
fractions <- read_tsv(system.file("extdata/study_design/fractions.txt", package = "PlexedPiperTestData"))
head(fractions)
```

## Samples
This links plex ID, reporter channel with samples. There is a minor difference
between ReporterAlias and MeasurementName. MeasurementName is the sample name
that will be finally reported. ReporterAlias is an itermediate between 
ReporterName and MeasurementName and is used for defining what the reference is.
```{r read_samples}
samples <- read_tsv(system.file("extdata/study_design/samples.txt", package = "PlexedPiperTestData"))
head(samples,10)
```

## References
Reference can be a certain channel, average of multiple channels or 1.
In general form it is an expression with ReporterAlias names as variables. It
is evaluated for each PlexID/QuantBlock combination and applied 
to divide reporter ion intensities within corresponding PlexID/QuantBlock.
QuantBlock can be though of as a way of defining sub-plex. In a typical TMT
experiment QuantBlock is always 1. In case of 5 pairwise comparisons within
TMT10, there will be 5 QuantBlocks (1..5) with reference for each QuantBlock.
```{r read_references}
references <- read_tsv(system.file("extdata/study_design/references.txt", package = "PlexedPiperTestData"))
head(references)
```


## Creating quantitative cross-tab
Final step when MS/MS IDs and reporter ions linked together, aggregated down to
peptide or accession (i.e. protein) level. To retain protein IDs while 
aggregating to peptide level we suggest 
`aggregation_level <- c("accession","peptide")`
```{r crosstab_compillation}
aggregation_level <- c("accession")
quant_cross_tab <- create_crosstab(msnid, 
                                   masic_data, 
                                   aggregation_level, 
                                   fractions, samples, references)
dim(quant_cross_tab)
head(quant_cross_tab)
```



```{r cleanup, echo=FALSE}
unlink(".Rcache", recursive=TRUE)
```
